Scalar Items in Embedded Position: {A}n Experimental Revisit

### 2012-08-03 text/paper/paper1.tex
beginning of prosody part revision

% We hypothesized that previous studies might be
% insufficiently informative because of the focus on (variants of) a
% picture-verification paradigm.\dn{relate to Bob van Tiel's work} The
% problem is that in order to test the availability of different
% candidate meanings different pictures have to be presented, so that
% effects of pictorial complexity or stereotypicality could never be
% ruled out entirely.
Moreover, previous studies have only accumulated limited evidence
pertaining to the second question that may help decide between
theoretical positions, namely which of the attested readings subjects
prefer. Finally, previous studies presented target sentences
visually. But as it is often argued that intonational stress on an
embedded scalar item can favor a local reading
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar},
it seems important to present target sentences auditorily, so as to be
able to control for effects of ``silent intonation'' that subjects may
apply when reading a sentence see Bader 1998, \citep{Fodor1998:Learning}
\dn{provide
  references}.

In reaction to this situation, we therefore presented sentence
materials auditorily and also employed a different kind of visual
presentation of the pictorial material.
%: subjects were presented with
%pictures that were initially covered and could incrementally be
%uncovered at the subjects' request; at each step of uncovering,
%subjects had to decide whether they could already give a truth-value
%judgement or needed more information (Comry ????)\dn{insert
%  references}. 
This way we obtained behavioral data that is both
indicative of the reading subjects assumed and independent of the
confounding pictorial effects. 
%At the same time, we hypothesized
%that the incremental nature of this task would shed light on the
%preferences over readings, because the temporal distribution of
%truth-value judgments would give away which reading subjects were
%waiting to evaluate, so to speak. 
Adding to previous studies, we made sure that our method is able 
to reveal information about reading preferences by relating our 
targets to strucutres that are known in the psycholinguistic 
literature for exhibiting certain preferences.

@@

\paragraph{Testing Effects of accentuation.}\dn{These two paragraphs
  still need a proper home; Petra opts for ``materials'' section.} It
has been claimed that local readings only emerge in the case of
special accentuation of the the scalar item under consideration
(REFERENCES). Further, it is generally assumed that the required type
of accentuation is a pitch accent (REFS). To test for this possibility
we decided to present the sentences (eg., (\ref{ex:as}) and
(\ref{ex:es})) auditorily in two versions each. In the accented
version a pitch accent was placed on {\it einige}. The second version
had neutral prosody. If accentuation is the driving force for local
readings, we would expect a higher proportion of local readings in the
former than in the latter version of the sentences.

>>>>

% We hypothesized that previous studies might be
% insufficiently informative because of the focus on (variants of) a
% picture-verification paradigm.\dn{relate to Bob van Tiel's work} The
% problem is that in order to test the availability of different
% candidate meanings different pictures have to be presented, so that
% effects of pictorial complexity or stereotypicality could never be
% ruled out entirely.
Moreover, previous studies have only accumulated limited evidence
pertaining to the second question that may help decide between
theoretical positions, namely which of the attested readings subjects
prefer. Finally, previous studies presented target sentences
visually. But as it is often argued that focal stress on an
embedded scalar item can favor a local reading
\citep[e.g.][]{Horn2006:The-Border-Wars,Geurts2009:Scalar-Implicat,ChemlaSpector2010:Experimental-Ev,Geurts2010:Quantity-Implic,Tielvan-Tiel2012:Embedded-Scalar},
it seems important to present target sentences auditorily, so as to be
able to control for effects of ``silent prosody'' that subjects may
apply when reading a sentence see Bader 1998, \citep{Fodor1998:Learning}
\dn{provide
  references}.

In reaction to this situation, we therefore presented sentence
materials auditorily and also employed a different kind of visual
presentation of the pictorial material.
%: subjects were presented with
%pictures that were initially covered and could incrementally be
%uncovered at the subjects' request; at each step of uncovering,
%subjects had to decide whether they could already give a truth-value
%judgement or needed more information (Comry ????)\dn{insert
%  references}. 
This way we obtained behavioral data that is both
indicative of the reading subjects assumed and independent of the
potential pictorial and prosodic confounds. 
%At the same time, we hypothesized
%that the incremental nature of this task would shed light on the
%preferences over readings, because the temporal distribution of
%truth-value judgments would give away which reading subjects were
%waiting to evaluate, so to speak. 
Adding to previous studies, we made sure that our method is able 
to reveal information about reading preferences by relating our 
targets to strucutres that are known in the psycholinguistic 
literature for exhibiting certain preferences.

@@

\paragraph{Testing Effects of prosody.}\dn{These two paragraphs
  still need a proper home; Petra opts for ``materials'' section.}

The availability of local readings has often been claimed to be
affected by the presence or absence of a focal accent on the scalar
item. By contrastively accenting the scalar, potential alternatives
are highlighted, resulting in more implicatures being computed.
Some approaches have even claimed that local readings exclusively 
emerge when the scalar item is accented and are otherwise generally
unavailable (Refs).

Though there seems to be general consensus about these potential 
prosodic effects, empirical evidence for this assumption 
is still scarce. One notable exception comes from a pilot study 
by Frazier, in which participants were instructed to read embedded
and non-embedded versions of English ae sentences, in which some was either
capitalized or non-capitalized. Prior to the experiment, participants 
were told that capitalizing corresponded to contrastive stress.
Interestingly, strengthened interpretations occurred rather often 
in the embedded sentences in this task (59 \%), but were 
not affected by capitalizing. ...
Another study relevant for the present consideration comes from Schwarz et al., 
examining the effects of contrastive accentuation of "or" in non-embedded sentences 
like {\it Mary will invite Fred or/OR Sam to barbecue}. After listening to these 
sentences, participants were presented two alternative readings ({\it 1. She will 
invite Fred or Sam or possibly both}, or {\it 2. She will invite Fred or Sam but 
not both}).

%  It
%has been claimed that local readings only emerge in the case of
%special accentuation of the the scalar item under consideration
%(REFERENCES). Further, it is generally assumed that the required type
%of accentuation is a pitch accent (REFS). To test for this possibility
%we decided to present the sentences (eg., (\ref{ex:as}) and
%(\ref{ex:es})) auditorily in two versions each. In the accented
%version a pitch accent was placed on {\it einige}. The second version
%had neutral prosody. If accentuation is the driving force for local
%readings, we would expect a higher proportion of local readings in the
%former than in the latter version of the sentences.
